---
title: "**Industrial Use**"

author: |
  | **Bruno Caetano Vidigal**
  | Food and Agriculture Organization of the United Nations
output: pdf_document
---
---

```{r include=FALSE, cache=FALSE}
## Load required functions
library(data.table)
library(ggplot2)
library(scales)
library(knitr)
library(faosws)
library(faoswsUtil)
library(faoswsFlag)
library(magrittr)
library(igraph)
library(dplyr)

```

```{r include=FALSE, cache=FALSE}
library(data.table)
library(countrycode)

## eurostat data: biodiesel

setwd("~/Github/faoswsIndustrial/Data")

# ratio: the piece of information about "ethanol" came from ebb and for "biodiesel" 
# from one website (http://www.paulchefurka.ca/Biofuels.html).
ratioAdditionalSupplyBiofuelDemand = fread("ratioAdditionalSupplyBiofuelDemand.csv")
ratioAdditionalSupplyBiofuelDemand

# structure : this dataset came from ebb
biofuelStructure = fread("biofuelStructure.csv")
biofuelStructure

# Biodiesel from eurostat
eurostatBiodieselTOE = fread("eurostatBiodieselTOE.csv")
head(eurostatBiodieselTOE)

eurostatBiodieselTOE[, m49 := countrycode(country, "country.name", "un")]
eurostatBiodieselTOE = eurostatBiodieselTOE[!is.na(m49)]

eurostatBiodieselTOE

# melt
eurostatBiodieselTOE = melt.data.table(eurostatBiodieselTOE, 
                                       id.vars = c("country", "m49"), 
                                       measure.vars = paste0("year", 2003:2014))

eurostatBiodieselTOE = eurostatBiodieselTOE[!(value == ":")]
eurostatBiodieselTOE[, year := substr(variable, 5, 8)]
eurostatBiodieselTOE[, c("variable") := NULL]
setcolorder(eurostatBiodieselTOE, c("country", "m49", "year", "value"))
eurostatBiodieselTOE[, biofuel := "biodiesel"]

# Biogasoline from eurostat
eurostatBiogasolineTOE = fread("eurostatBiogasolineTOE.csv")
head(eurostatBiogasolineTOE)

eurostatBiogasolineTOE[, m49 := countrycode(country, "country.name", "un")]
eurostatBiogasolineTOE = eurostatBiogasolineTOE[!is.na(m49)]

eurostatBiogasolineTOE

# melt
eurostatBiogasolineTOE = melt.data.table(eurostatBiogasolineTOE, 
                                       id.vars = c("country", "m49"), 
                                       measure.vars = paste0("year", 2003:2014))

eurostatBiogasolineTOE = eurostatBiogasolineTOE[!(value == ":")]
eurostatBiogasolineTOE[, year := substr(variable, 5, 8)]
eurostatBiogasolineTOE[, c("variable") := NULL]
setcolorder(eurostatBiogasolineTOE, c("country", "m49", "year", "value"))
eurostatBiogasolineTOE[, biofuel := "bioethanol"]

## join both biodiesel and biogasoline data

euroStatBiofuel = rbind(eurostatBiodieselTOE, eurostatBiogasolineTOE)
```

# Abstract

This document provides detailed description of how we are using the data sets **biodiesel** and **biogasoline** from **EuroStat** for spliting the figures into the commodity level to the countries of European Union. We are considering that **biogasoline** is **bioethanol**.

# Approach 1

In this approach we are using the share's table from **European Biodiesel Board (EBB)** for spliting the production of **biofuels** for **country/commodity** level.

## 1.1 Sources of data sets

We are pulling the data sets **biodiesel** and **biogasoline** between **2003** and **2014** from the **EuroStat**:

http://ec.europa.eu/eurostat/tgm/refreshTableAction.do?tab=table&plugin=1&pcode=ten00081&language=en

After combine both, we get this data set below. In this example, France produced 1.618 (**1000  tonnes of oil equivalent**) of biodiesel and 423,7 (**1000  tonnes of oil equivalent**) of bioethanol. 

````{r, echo=FALSE}
bioethanol = head(euroStatBiofuel[biofuel == "bioethanol" & country == "France" & year == 2011], 3)
biodiesel = head(euroStatBiofuel[biofuel == "biodiesel" & country == "France" & year == 2011], 3)

kable(rbind(biodiesel, bioethanol), format = "markdown", padding = 0, 
      col.names=c("country", "m49", "year", "value", "biofuel"))
````

This piece of information from EuroStat is the primary production of **biodiesel** and **bioethanol** just for the **country level** and we need to figure out **how many tonnes of each commodity** was used to produce biofuel. In order to do that we pulled a piece of information from others sources.
First, we pulled the Conversion Efficiency (Litres/tonne) from **FAO** (http://www.greenfacts.org/en/biofuels/figtableboxes/biofuel-yields-countries.htm) and from another website (http://www.paulchefurka.ca/Biofuels.html).
From this table we can see that 1 tonne of **Sugar Beet** produces 110 litres of biodiesel.

````{r, echo=FALSE}
kable(ratioAdditionalSupplyBiofuelDemand, format = "markdown", padding = 0, 
      col.names=c("commodity", "biofuel", "litres2tonne"))
````

We also pulled the average proportion of each commodity used to produce **biodiesel** and **bioethanol** from **European Biodiesel Board (EBB)** and the density of each commodity from FAO (http://www.fao.org/docrep/017/ap815e/ap815e.pdf). So, **53.1%** of **biodiesel** came from **Rapeseed** and its density is **0,92g/ml**. Also we can say that 4.7% of all **biodiesel** was produced from Sunflower.

````{r, echo=FALSE}
biofuelStructure[, refScenario2 := 100 * round(refScenario, 3)]
kable(biofuelStructure[, c("commodity", "biofuel", "timePointYears", "refScenario2", "densityG_ml"), with = F], format = "markdown", padding = 0, 
      col.names=c("commodity", "biofuel", "year", "%(EBB)", "density"))
````

## 1.2 Spliting production by commodity 

We merged the data sets to figure out how many tonnes were produced from different commodities based on EBB's table. So, in the table below you find the **valueByCommodityTonnes** what means the **amount of biofuel** produced from the specific commodity. For example, we can say that 169,953 tonnes of biodiesel came from **Palm Fruit** in France in 2011.

```{r include=FALSE, cache=FALSE}

# merge with biofuelStructure
euroStatBiofuel = merge(euroStatBiofuel, biofuelStructure[, c("commodity",
                                                              "refScenario", 
                                                              "biofuel", 
                                                              "densityG_ml"),
                                                          with = F], by = "biofuel",
                        all.x=T, allow.cartesian=TRUE)

euroStatBiofuel[, value := as.numeric(value)]
euroStatBiofuel[, valueByCommodityTonnes := value * refScenario * 1000]

# merge eurostatBiodieselTOE with the conversion factors
euroStatBiofuel = merge(euroStatBiofuel, ratioAdditionalSupplyBiofuelDemand[, c("commodity",
                                        "conversion_efficiency_litres_tonne"),
                                        with = F], by = "commodity", all.x=T)

````


````{r, echo=FALSE}
bioethanol = euroStatBiofuel[biofuel == "bioethanol" & country == "France" & year == 2011, c("country", "biofuel", "commodity", "year", "value", "refScenario", "densityG_ml", "conversion_efficiency_litres_tonne", "valueByCommodityTonnes"), with=F]

biodiesel = euroStatBiofuel[biofuel == "biodiesel" & country == "France" & year == 2011, c("country", "biofuel", "commodity", "year", "value", "refScenario", "densityG_ml", "conversion_efficiency_litres_tonne", "valueByCommodityTonnes"), with=F]

biofuelRbind = rbind(biodiesel, bioethanol)
biofuelRbind[, refScenario2 := 100 * round(refScenario, 3)]

kable(biofuelRbind[, c("country", "biofuel", "commodity", "year", "value", "refScenario2", "densityG_ml", "conversion_efficiency_litres_tonne", "valueByCommodityTonnes"), with=F], format = "markdown", padding = 0, 
      col.names=c("country", "biofuel", "commodity", "year", "totalValue", "%(EBB)", "density", "litres2tonne", "valueByCommodityTonnes"))
````

# 1.3 Calculating how many tonnes of commodity was used 

We can calculate how many tonnes of each commodity was used to produce the amount above. In order to do that, we have to convert from **litres** to **tonnes** using the **density**. So, we know for example that **212,176.03 tonnes** of **Palm Fruit** was used to produce **1,618.6** (**1000  tonnes of oil equivalent**).
````{r, echo=FALSE}
# we need to convert from litres to tonnes

euroStatBiofuel[, conversion_litres_tonne := (densityG_ml * conversion_efficiency_litres_tonne)/1000]

## Now we can calculate how many tonnes of each commodity was used
euroStatBiofuel[, prodByCommodity := valueByCommodityTonnes/conversion_litres_tonne]
````


````{r, echo=FALSE}
bioethanol = euroStatBiofuel[biofuel == "bioethanol" & country == "France" & year == 2011, c("country", "biofuel", "commodity", "year", "value", "refScenario", "densityG_ml", "prodByCommodity"), with=F]

biodiesel = euroStatBiofuel[biofuel == "biodiesel" & country == "France" & year == 2011, c("country", "biofuel", "commodity", "year", "value", "refScenario", "densityG_ml", "prodByCommodity"), with=F]

kable(rbind(biodiesel, bioethanol), format = "markdown", padding = 0, 
      col.names=c("country", "biofuel", "commodity", "year", "totalValue", "%(EBB)", "density", "prodByCommodity"))
````

# Approach 2

We also could use the Production and Net Trade to split by **country/commodity level** the amount of **biodiesel** and **bioethanol** produced. Let's do that only for **biodiesel**. We'll pull the data from de old system for 2011 with the variables: Production, Imports, Stock variation, Exports, Feed, Food and Other Util for the commodities **Oil palm fruit**, **Palm oil**, **Oil of Palm Kernel**, **Cake of  Palm kernel**, **Rapeseed or canola oil, crude**, **Cake of Rapeseed**, **Soya bean oil**, **Cake of  Soya beans**, **Sunflower-seed oil, crude** and **Cake of  Sunflower seed**.

```{r include=FALSE, cache=FALSE}

if(!exists("DEBUG_MODE") || DEBUG_MODE == "") {
  if(Sys.info()[7] == "josh"){ # Josh's work computer
    R_SWS_SHARE_PATH <- "/media/hqlprsws1_qa/"
    SetClientFiles(dir = "~/R certificate files/QA/")
    files = dir("~/Documents/Github/faoswsFood/R",
                full.names = TRUE)
    token = "557f0e65-5f84-43b0-a021-fe3bf3f02316"
  } else if(Sys.info()[7] == "caetano"){ # bruno's work computer
    SetClientFiles(dir = "~/.R/QA/")
    R_SWS_SHARE_PATH = "//hqlprsws1.hq.un.fao.org/sws_r_share"
    files = dir("~/Github/faoswsFood/R", full.names = TRUE)
    token = "66a36f31-1a29-4a49-8626-ae62117c251a"
  } else {
    stop("User not yet implemented!")
  }  
  
  GetTestEnvironment(
    baseUrl = "https://hqlqasws1.hq.un.fao.org:8181/sws",
    token = token
  )
 }

## This code below is pulling data from the old system for "Other util"

# 51 production (Mt)
# 61 imports (Mt)
# 71 stock variation (Mt) 
# 91 exports (Mt)  
# 101 feed (Mt)
# 131 processed (Mt)
# 141 food (Mt)
# 151 other util

elements = c("51", "61", "71", "91", "101", "131", "141", "151")

# As we want to analyse just the results for biodiesel, let's pull just the 
# commodities used to produce it. The names of the commodities are: "Oil palm fruit",
# "Palm oil", "Oil of Palm Kernel", "Cake of  Palm kernel", "Rapeseed or canola oil, crude",
# "Cake of  Rapeseed", "Soya bean oil", "Cake of  Soya beans", "Sunflower-seed oil, crude",
# "Cake of  Sunflower seed"

items = c("254", "257", "258", "259", "271", "272", "237", "238", "268", "269")

key = DatasetKey(domain = "faostat_one", dataset = "FS1_SUA_UPD", dimensions = list(
  Dimension(name = "geographicAreaFS", keys = GetCodeList("faostat_one", "FS1_SUA_UPD", "geographicAreaFS")[, code]),
  Dimension(name = "measuredElementFS", keys = elements),
  Dimension(name = "timePointYears", keys = as.character(2011)),
  Dimension(name = "measuredItemFS", keys = items))
)
data = GetData(key)

# dcast.data.table
data <- dcast.data.table(data,
                         geographicAreaFS + measuredItemFS + timePointYears ~ measuredElementFS,
                         value.var = "Value")


setnames(data, old=c("101", "131", "141", "151", "51", "61", "71", "91"),
         new=c("feed", "processed", "food", "otherUtil", "production", "imports", "stockVariation", "exports"))


data[, geographicAreaM49 := fs2m49(as.character(geographicAreaFS))]

data[, measuredItemFS := formatC(as.numeric(measuredItemFS), width = 4,
                                 flag = "0")]
data[, measuredItemCPC := fcl2cpc(as.character(measuredItemFS))]
data

# names of the commodities and countries
data <- nameData("agriculture", "aproduction", data)

data[grepl("palm", tolower(measuredItemCPC_description))][, .N, c("measuredItemCPC_description", "measuredItemFS")]
data[grepl("palm", tolower(measuredItemCPC_description)), commodity := "Palm Fruit"]


data[grepl("rapeseed", tolower(measuredItemCPC_description))][, .N, c("measuredItemCPC_description", "measuredItemFS")]
data[grepl("rapeseed", tolower(measuredItemCPC_description)), commodity := "Rapeseed"]


data[grepl("soya", tolower(measuredItemCPC_description))][, .N, c("measuredItemCPC_description", "measuredItemFS")]
data[grepl("soya", tolower(measuredItemCPC_description)), commodity := "Soybeans"]

data[grepl("sunflower", tolower(measuredItemCPC_description))][, .N, c("measuredItemCPC_description", "measuredItemFS")]
data[grepl("sunflower", tolower(measuredItemCPC_description)), commodity := "Sunflower"]

# no NA's
for(cname in c("feed", "processed", "food", "otherUtil", "production", 
               "imports", "stockVariation", "exports")){
  data[is.na(get(cname)), c(cname) := 0]
}


data1 = data[, list(production = sum(production),
                    imports = sum(imports),
                    exports = sum(exports),
                    stockVariation = sum(stockVariation)), 
             by=list(geographicAreaM49_description, geographicAreaM49, timePointYears, measuredItemCPC,commodity)]

data1[, prodNetTrade := production + imports - exports - stockVariation]
data1[, geographicAreaM49:= as.factor(geographicAreaM49)]
````


## 2.1 Calculating the food availability

The food availability is Production + Imports - Exports - Stock_variation. So, in this case we have 614,162 tonnes of **Soybeans** available in France in 2011.
````{r, echo=FALSE}

kable(data1[geographicAreaM49_description == "France" & timePointYears == 2011, c("geographicAreaM49_description", "commodity", "timePointYears", "production", "imports", "exports", "stockVariation", "prodNetTrade"), with=F], format = "markdown", padding = 0, 
      col.names=c("country", "commodity", "year", "production", "imports", "exports", "stockVariation", "prodNetTrade"))

````


```{r include=FALSE, cache=FALSE}
# let's read the table with the flag to EU28
setwd("C:/Users/caetano/Documents/Github/faoswsIndustrial/sandbox")
europeanCountryCode = fread("europeanCountryCode.csv")
europeanCountryCode[, geographicAreaM49 :=  as.factor(countrycode(countryName, "country.name", "un"))]
europeanCountryCode = europeanCountryCode[flagEU28 == 1]

# merge data with the the flag of EU28
data1 = merge(data1, europeanCountryCode[, c("geographicAreaM49", "flagEU28"), with = F],
      by = "geographicAreaM49", all.x=T)

data1[is.na(flagEU28), flagEU28 := 0]
data1[, flagEU28 := as.factor(flagEU28)]

# by country, flagEU28 and commodity
countryByCommodity = data1[, list(prodNetTrade = sum(prodNetTrade)),
     by=list(geographicAreaM49, geographicAreaM49_description, flagEU28, commodity)]

countryByCommodity[, biofuelStructByCountry := prodNetTrade/sum(prodNetTrade), 
     by = geographicAreaM49]

# by flagEU28 and commodity
flagEU28Commodity = data1[, list(prodNetTrade = sum(prodNetTrade)),
     by=list(flagEU28, commodity)]

flagEU28Commodity[, biofuelStructEU28 := prodNetTrade/sum(prodNetTrade), 
     by = flagEU28]

# Now, we have to merge the table countryByCommodity with  the table 
# euroStatBiofuel.

keys = c("geographicAreaM49", "commodity")

setnames(euroStatBiofuel, "m49", "geographicAreaM49")
euroStatBiofuel[, geographicAreaM49 := as.factor(geographicAreaM49)]

euroStatBiofuel = merge(euroStatBiofuel, countryByCommodity[, c("geographicAreaM49", "commodity",
                                              "biofuelStructByCountry"), with = F], by=keys, all.x=T)



````

## 2.2 Merge tables

Now we can merge this data set with the data set previous with the **density** and **conversion efficiency** information. We see that 39.9% of **biodiesel** came from **Rapeseed**.

````{r, echo=FALSE}

euroStatBiofuel[, biofuelStructByCountry2 := 100*round(biofuelStructByCountry, 3)]

kable(euroStatBiofuel[country == "France" & biofuel == "biodiesel" & year == 2011, c("country", "commodity", "year", "densityG_ml", "conversion_efficiency_litres_tonne", "biofuelStructByCountry2"), with=F], format = "markdown", padding = 0, 
      col.names=c("country", "commodity", "year", "density", "conversion_efficiency_litres_tonne", "% ProdNetTrade"))

````


```{r include=FALSE, cache=FALSE}

euroStatBiofuel[biofuel == "biodiesel", valueBiodiselByCommodityTonnes := value * biofuelStructByCountry * 1000]

# we need to convert from litres to tonnes

## Now we can calculate how many tonnes of each commodity was used
euroStatBiofuel[biofuel == "biodiesel", prodBiodiselByCommodity := valueBiodiselByCommodityTonnes/conversion_efficiency_litres_tonne]
euroStatBiofuel
````

## 2.3 Calculating how many tonnes of commodity was used

Using the same rules applied before, we get the result below. In this case
63.81488 tonnes of Palm Fruit was used to produce 1,618.6 (1000 tonnes of oil equivalent).

````{r, echo=FALSE}

kable(euroStatBiofuel[country == "France" & biofuel == "biodiesel" & year == 2011, c("country", "commodity", "year", "value", "biofuelStructByCountry2","densityG_ml", "prodBiodiselByCommodity"), with=F], format = "markdown", padding = 0, 
      col.names=c("country", "commodity", "year", "totalValue", "% ProdNetTrade", "density", "prodByCommodity"))

````
