---
title: "**Industrial Use**"

author: |
  | **Bruno Caetano Vidigal**
  | Food and Agriculture Organization of the United Nations
output: pdf_document
---
---

```{r include=FALSE, cache=FALSE}
## Load required functions
library(data.table)
library(ggplot2)
library(scales)
library(knitr)
library(faosws)
library(faoswsUtil)
library(faoswsFlag)
library(magrittr)
library(igraph)
library(dplyr)

```

```{r include=FALSE, cache=FALSE}
DEBUG_MODE = Sys.getenv("R_DEBUG_MODE")
R_SWS_SHARE_PATH = Sys.getenv("R_SWS_SHARE_PATH")

print_cs1 <- function(d) cat(paste(shQuote(d, type="cmd"), collapse=", "))

if(!exists("DEBUG_MODE") || DEBUG_MODE == ""){
  token = "41558a20-c419-4821-8288-2dc7ccbc5ecf"
  GetTestEnvironment("https://hqlqasws1.hq.un.fao.org:8181/sws", token)
  
}

source("~/Github/faoswsIndustrial/R/getBioFuelData.R")
source("~/Github/faoswsIndustrial/R/getCPCTreeItem.R")
source("~/Github/faoswsIndustrial/R/getCountryCodeSUA.R")
source("~/Github/faoswsIndustrial/R/getItemCommSUA.R")


## Extracting data from USDA domain/dataset

countryCodeInd <- GetCodeList("usda", "usda_psd", "geographicUsda")
elementInd <- GetCodeList("usda", "usda_psd", dimension = "measuredElementPsd")
itemInd <- GetCodeList("usda", "usda_psd", dimension = "measuredItemPsd")
yearRange <- as.character(1961:2015)

countryDim1 <- Dimension(name = "geographicUsda", 
                         keys = countryCodeInd[, code])

elementDim2 <- Dimension(name = "measuredElementPsd", 
                         keys = "140.08")

itemDim3 <- Dimension(name = "measuredItemPsd",
                      keys = itemInd[, code])

timePointYearsDim4 <- Dimension(name = "timePointYears",
                                keys = yearRange)

dataKey <- DatasetKey(domain = "usda", dataset = "usda_psd", 
                      dimensions = list(countryDim1, elementDim2, itemDim3, 
                                        timePointYearsDim4))

vegetableOilsDataForIndUses <- GetData(dataKey, flags = FALSE)
vegetableOilsDataForIndUses[, measuredElementPsd := NULL]
vegetableOilsDataForIndUses[, geographicAreaM49 := as.character(getCountryCodeSUA(geographicUsda))]
vegetableOilsDataForIndUses[, measuredItemCPC := as.character(getItemCommSUA(measuredItemPsd))]
vegetableOilsDataForIndUses[, c("geographicUsda", "measuredItemPsd") := NULL]

setcolorder(vegetableOilsDataForIndUses, c("timePointYears", "geographicAreaM49", "measuredItemCPC", "Value"))

## Pull agricFeedStuffsForBioFuelData
allCPCItem = getCPCTreeItem()
agricFeedStuffsForBioFuelData <- getBioFuelData()


agricFeedStuffsForBioFuelData$geographicAreaM49 <- as.character(agricFeedStuffsForBioFuelData$geographicAreaM49)
agricFeedStuffsForBioFuelData$measuredItemCPC <- as.character(agricFeedStuffsForBioFuelData$measuredItemCPC)
agricFeedStuffsForBioFuelData$timePointYears <- as.character(agricFeedStuffsForBioFuelData$timePointYears)

## Merge the two datasets

industrialUsesData <- merge(agricFeedStuffsForBioFuelData, vegetableOilsDataForIndUses,
                            by = c("geographicAreaM49", "measuredItemCPC", "timePointYears"),
                            all = TRUE)


industrialUsesData$Value_measuredElement_5150 <- as.numeric(industrialUsesData$Value_measuredElement_5150)
industrialUsesData$Value_measuredElement_5150[which(is.na(industrialUsesData$Value_measuredElement_5150))] <- 0
industrialUsesData$Value[which(is.na(industrialUsesData$Value))] <- 0

# Value_measuredElement_5150 is in tonns
# Value is in 1000MT

industrialUsesData[, Value_measuredElement_ind := Value_measuredElement_5150 + Value * 1000]

industrialUsesData[, c("Value", "Value_measuredElement_5150") := NULL]

industrialUsesData[, industrialElement:= "5195"]

setcolorder(industrialUsesData,
            c("timePointYears", "geographicAreaM49", "measuredItemCPC", "industrialElement",
              "Value_measuredElement_ind", "flagObservationStatus_measuredElement_5150", 
              "flagMethod_measuredElement_5150"))


setnames(industrialUsesData,
         old = c("timePointYears", "geographicAreaM49", "measuredItemCPC", "industrialElement",
                 "Value_measuredElement_ind", "flagObservationStatus_measuredElement_5150", 
                 "flagMethod_measuredElement_5150"),
         new = c("timePointYears", "geographicAreaM49", "measuredItemCPC", "measuredElement",
                 "Value", "flagObservationStatus", "flagMethod"))


industrialUsesData[, flagObservationStatus:= "I"]
industrialUsesData[, flagMethod:= "e"]

cpcCodes = GetCodeList("agriculture", "aproduction", "measuredItemCPC")
setnames(cpcCodes, "code", "measuredItemCPC")

countryNames = GetCodeList("agriculture", "aproduction", "geographicAreaM49")
setnames(countryNames, old = c("code", "description"), new=c("geographicAreaM49", "countryName"))

industrialUsesData <- merge(industrialUsesData, cpcCodes[, c("measuredItemCPC", "description"), with=F], by = "measuredItemCPC")
industrialUsesData <- merge(industrialUsesData, countryNames[, c("geographicAreaM49", "countryName"), with=F], by = "geographicAreaM49")

space <- function(x, ...) { 
  format(x, ..., big.mark = ".", scientific = FALSE, trim = TRUE)
}

```

# About this document

In this document there are some statistics and summaries about the **Industrial Use**.

# 1. Summaries

````{r, echo=FALSE}

tabSummary <- industrialUsesData[, list(country = length(unique(geographicAreaM49)),
                          commodity = length(unique(description)),
                          DateMin = min(timePointYears),
                          DateMax = max(timePointYears))]

kable(tabSummary, format = "markdown", padding = 0, 
      col.names=c("Number of distinct countries", "Number of distinct commodities", "Min Date", "Max Date"))

````

In the plot below, we can see the number of distinct countries with values for industrial utilization per year.
````{r, echo=FALSE}
tabYearCountry <- industrialUsesData[, list(value = sum(Value)),
                                 by=list(countryName, timePointYears)]

tabYearCountry <- tabYearCountry[order(-timePointYears, -value)]
tabYearCountry <- tabYearCountry[order(-timePointYears, -value)]
tabYearCountry[, percent := 100*round(value/sum(value), 3), by=timePointYears]
tabYearCountry[, cumsum := cumsum(percent), by=timePointYears]


tsCountryYear <- tabYearCountry[value > 0, .N, timePointYears]
tsCountryYear$timePointYears <- as.numeric(tsCountryYear$timePointYears)


ggplot(data=tsCountryYear, aes(x=timePointYears, y=N, group=1)) + 
  geom_line(colour="blue", size=1.2) + scale_y_continuous(lim=c(0, 80), breaks=seq(0, 80, 10)) + 
  scale_x_continuous(lim=c(1961, 2015),  breaks=seq(1961, 2015, 3)) + 
  xlab("Year") + ylab("Number of countries") + 
  theme(axis.text.x = element_text(angle = 45, vjust = .5))

````


# 2. Countries with more production for Industrial utilization in **2015**

Look at the 5 countries with more production for Industrial utilization represent  **`r paste0(max(head(tabYearCountry[, c("countryName", "value", "percent", "cumsum"), with=F], 5)$cumsum), '%')`** of all in 2015.
````{r, echo=FALSE}
options(warn=-1)
tabYearCountry[, value2 := sub(".000000", "", space(value))]
kable(head(tabYearCountry[, c("countryName", "value2", "percent", "cumsum"), with=F], 5), format = "markdown", padding = 0,
      col.names=c("Country", "Value", "%", "Cumulative %"))

````

# 3. Commodities with more production for Industrial utilization in **2015**

````{r, echo=FALSE}

tabYearCommodity <- industrialUsesData[, list(value = sum(Value)),
                                     by=list(description, timePointYears)]

tabYearCommodity <- tabYearCommodity[order(-timePointYears, -value)]

tabYearCommodity[, percent := 100*round(value/sum(value), 3), by=timePointYears]
tabYearCommodity[, cumsum := cumsum(percent), by=timePointYears]

tabYearCommodity[, value2 := sub(".000000", "", space(value))]

````

Look at the 5 commodities with more production for Industrial utilization represent  **`r paste0(max(head(tabYearCommodity[, c("description", "value", "percent", "cumsum"), with=F], 5)$cumsum), '%')`** of all in 2015.

````{r, echo=FALSE}

kable(head(tabYearCommodity[, c("description", "value2", "percent", "cumsum"), with=F], 5), format = "markdown", padding = 0,
      col.names=c("Commodity", "Value", "%", "Cumulative %"))

````

The plot below shows the number of commodities with industrial utilization.

````{r, echo=FALSE}
tsCommodityYear <- tabYearCommodity[value > 0, .N, timePointYears]
tsCommodityYear$timePointYears <- as.numeric(tsCommodityYear$timePointYears)


ggplot(data=tsCommodityYear, aes(x=timePointYears, y=N, group=1)) + 
  geom_line(colour="blue", size=1.2) + scale_y_continuous(lim=c(0, 20), breaks=seq(0, 20, 5)) + 
  scale_x_continuous(lim=c(1961, 2015),  breaks=seq(1961, 2015, 3)) + 
  xlab("Year") + ylab("Number of commodities") + 
  theme(axis.text.x = element_text(angle = 45, vjust = .5))
````

#4. Times series for the five countries with more industrial utilization in 2015

The plot below shows the times series for the five countries with more industrial utilization in 2015.
````{r, echo=FALSE}
top5countries2015 <- tabYearCountry[countryName %in% c("Brazil", "Indonesia", "Argentina", "China", "Malaysia")] 
top5countries2015$timePointYears <- as.numeric(top5countries2015$timePointYears)
top5countries2015[, value := value/1000]

#options(scipen=999)
ggplot(data=top5countries2015, aes(x=timePointYears, y=value, group=countryName, color=countryName)) + 
  geom_line(size=1.2) + 
  scale_y_continuous(lim=c(0, 7000), breaks=seq(0, 7000, 1000), labels = space) + 
  scale_x_continuous(lim=c(1961, 2015),  breaks=seq(1961, 2015, 3)) + 
  xlab("Year") + ylab("Quantity (M)") +
theme(legend.title=element_blank()) + 
  theme(axis.text.x = element_text(angle = 45, vjust = .5))
````


#5. Times series for the five countries with more industrial utilization in 2015

The plot below shows the times series for the five commodities with more industrial utilization in 2015.
````{r, echo=FALSE}
top5Commodities2015 <- tabYearCommodity[description %in% c("Palm oil", "Soya bean oil", "Oil of Palm Kernel", 
                                                         "Flours and meals of oil seeds or oleaginous fruits, except those of mustard", 
                                                         "Sunflower-seed and safflower-seed oil")] 
top5Commodities2015$timePointYears <- as.numeric(top5Commodities2015$timePointYears)
top5Commodities2015[, value := value/1000]

top5Commodities2015[, description2 := ifelse(description == "Flours and meals of oil seeds or oleaginous fruits, except those of mustard", 
                                           "Flours and meals of oil seeds\n or oleaginous fruits, \nexcept those of mustard",
                                           ifelse(description == "Sunflower-seed and safflower-seed oil", "Sunflower-seed and \n safflower-seed oil",
                                           as.character(description)))]



#options(scipen=999)
ggplot(data=top5Commodities2015, aes(x=timePointYears, y=value, group=description2, color=description2)) + 
  geom_line(size=1.2) + 
  scale_y_continuous(lim=c(0, 15000), breaks=seq(0, 15000, 1000), labels = space) + 
  scale_x_continuous(lim=c(1961, 2015),  breaks=seq(1961, 2015, 3)) + 
  xlab("Year") + ylab("Quantity (M)") +
  theme(legend.title=element_blank()) + 
  theme(axis.text.x = element_text(angle = 45, vjust = .5))
````